\chapter{Acknowledgements}
First of all I want to thank Katja Mombaur for letting me work at her group Optimization, Robotics and Biomechanics. Being a part of her group has been very insightful and enjoyable throughout the time that I spent there. It was an honour for me to be given the chance and work on humanoid robots, to which not many have access to.\\\\
Secondly, I want to thank Kevin Stein, Yue Hu, and Manuel Kudruss for always lending me a hand, and for supporting the creation of this thesis. I want to further thank Matthew Millard, Giorgos Marinou, Jonas Große Sundrup, Julian Martus, Silvan Lindner, Alex Schubert, and Benjamin Reh for their uncountable amount of advices that they gave me on several topics, including but not limited to bug-fixing, structuring of this thesis and on general questions. Additionally, I want to thank Maximilian Staiger for his implementation of the android joystick app, and his very cooperative way of working, when we linked our implementations. Moreover, I want to thank Patrick Leibersperger, Christoph Rieß, and Lucas Möller for helping me to perform the data acquisition with Heicub. Especially, I want to thank Lucas Möller for providing me with access to powerful GPU clusters, which enabled me to train the neural networks in a reasonable amount of time. Once more, I want to thank Christoph Rieß for proofreading this thesis. Although it is not directly related to this thesis, I want to thank Markus Miltner for letting me stay at his place for the last couple of month, without whom I may have not been able to finish the work.\\\\
Finally, I want to thank my parents for their strength and their unconditional support that allowed me to study physics, and to follow my personal aims.